---
title: "607Assignment3B Dylan Gold"
format: html
editor: visual
---

# 607 Assignment 3B Dylan Gold

## Approach

In this assignment we have to work with a data set that has time series. We have to find/generate a data set then use sql or dplyr to calculate the average from year to year, as well as every 6 days I will find something online, maybe ask a LLM to generate one or give me ideas.

I will likely get the data as a csv onto github then use dplyr to create these averages after importing the data into R

## Codebase:

I started off by asking a LLM, (GPT 4) to give me a data set that satisfied the conditions for this assignment. The LLM conversation is in the LLM_Conversation3B.txt file.
It asked me to be a bit more specific on what kind of data I wanted. Due to the price of electricity going a lot these days, from what I have heard in large part due to data centers being built. I was curious about this so I told the LLM to give me some data on this.

The LLM prompted me to go to www.eia.gov/electricity/wholesalemarkets/csv/nyiso_load_act_hr_YYYY.csv
By changing the YYYY to a specific year we could get different years.

The LLM suggested I download the years I wanted then condense it into daily information. This is what I did. I got data from 2022 to 2025.
Because the file size is large, I decided to just download directly without importing it into github.

```{r , message=FALSE, warning=FALSE}
library(tidyverse)
library(ggplot2)
library(gt)

```

Get the 2022 Data
```{r, warning=FALSE}
url <- "http://www.eia.gov/electricity/wholesalemarkets/csv/nyiso_load_act_hr_2022.csv"

df <- read_csv(
  file = url,
  show_col_types = FALSE
)

head(df, 10)
```